---
title: "An Introduction to R"
author: "Daniel Kaplan"
date: "March 3, 2015"
output: ioslides_presentation
runtime: shiny
---

## R for Python People

```{r engine='python'}
print("Hello, Data^3 Python People!")
```

```{r}
"And to you also, R Folks!"
```

I'm orienting this talk to people who are familiar with programming ... but not necessarily either R or Python.

## Outline

1. Basic syntax of R
2. The community of packages
    * vetted and unvetted channels
    * writing and distributing packages
3. Data- and graphics-oriented packages
    * simple data scraping
    * some transformation, e.g. joins
    * ggplot2
    * dygraphs
4. Integration with documents
    * html & markdown
5. Interactivity
    * interactive documents
    * apps

## Good or Bad?

A language ought to be evaluated based on its effectiveness for the purpose at hand.

* Can people write the programs they need?
* Does it facilitate using existing, appropriate tools?  
* 

## Shiny Presentation

This R Markdown presentation is made interactive using Shiny. The viewers of the presentation can change the assumptions underlying what's presented and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Slide with Interactive Plot

```{r, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```


