---
title: "Bike Sharing in DC"
author: "Daniel Kaplan"
date: "March 4, 2015"
output: html_document
runtime: shiny
---

```{r include=FALSE}
library("leafletR")
# Install leaflet package
if(!require(leaflet)) 
  devtools::install_github("rstudio/leaflet")
library(leaflet)
library(lubridate)
library(dplyr)
library(ggplot2)
library("dygraphs")
library("XML")
```

* Data from Capital BikeShare

    * [Rental events](http://www.capitalbikeshare.com/trip-history-data)
    * [Station info](https://www.capitalbikeshare.com/data/stations/bikeStations.xml).  Check to see if this is updated to get information on the number of bikes at each station at different times.

* General Data

    * [Shape file for DC]
  
## Cleaning and reorganizing the rental events 

```{r}
csv_name <- "2014-Q4-Trips-History-Data.csv"
rda_name <- gsub(".csv", ".rda", csv_name)
```

```{r eval=FALSE}
Bikes <- as.data.frame(data.table::fread(csv_name))
```

```{r eval=FALSE}
Bikes <- 
  Bikes %>%
  rename(sdate=`Start date`, edate=`End date`, sstation=`Start Station`, estation=`End Station`,
         bikeno=`Bike#`, client=`Subscription Type`, duration=Duration) %>%
  mutate(sdate=lubridate::ymd_hm(sdate),
         edate=lubridate::ymd_hm(edate))
save(Bikes, file = rda_name)
```

## Reading the station info

```{r}
station_file <- "bikeStations.xml"
xml_data <- XML::xmlParse(file = station_file) %>% (XML::xmlToList)
Stations <- sapply(xml_data, unlist) %>% as.data.frame(stringsAsFactors = FALSE) %>% t %>% 
  as.data.frame(stringsAsFactors = FALSE)
Stations <- Stations[-nrow(Stations),]
Stations$lat  <- as.numeric(Stations$lat)
Stations$long <- as.numeric(Stations$long)
```

```{r}
load(rda_name)
```

## Task 1

A table of hourly use, weekdays and weekends.

```{r}
Bikes <- 
  Bikes %>%
  mutate(hour = lubridate::hour(sdate), 
         wday = lubridate::wday(sdate),
         weekend = ifelse( wday==1 | wday==7, "Weekend", "Midweek"),
         hour_in_week = hour + 24 * (wday - 1) )
Times <-
  Bikes %>% 
  group_by(hour, weekend)  %>%
  summarise(count=n())
```

```{r fig.height=3,fig.width=5}
ggplot(Times, aes(x=hour,y=count,group=weekend)) + 
  geom_line(aes(color=weekend)) 
```

```{r}
Bikes %>%
  group_by(hour_in_week)  %>%
  summarise(count=n()) %>% 
  ggplot( aes(x=hour_in_week, y=count)) +
  geom_line()
```

## Task 2

Density of use by hour for members and casuals

```{r}
Bikes %>%
  group_by(hour_in_week, client)  %>%
  summarise(count = n()) %>% 
  ggplot( aes(x = hour_in_week, y=count)) +
  geom_line(aes(color = client))
```

## Map of the stations

```{r}
Sdata <- 
  Stations %>% select(lat, long, nbBikes, nbEmptyDocks,name) 
renderLeaflet({
  leaflet(Sdata) %>% addTiles() %>% setView(-77.04, 38.9, zoom=12) %>%
    addCircleMarkers(radius=Sdata$nbBikes) %>%
    addCircleMarkers(radius=pmax(0,Sdata$nbEmptyDocks), color="red")
})
# renderLeaflet({
#   leaflet(dat, dest=tempdir(), size=c(800,600),center=c(38.9, -77.04))
#   })
```

```{r eval=FALSE}
sliderInput('n', 'Number of markers', min = 1, max = 100, value = 30, step = 1)
renderLeaflet({
  n = input$n
  df = data.frame(lat = rnorm(n, 42), lng = rnorm(n, -93))
  leaflet(df) %>% addTiles() %>% addMarkers()
})
```




## Time series

Smooth and display


## Task 3

Net influx and outflux by station, smoothed by seconds
```{r}
smooth_scale <- 300 # seconds
Bikes2 <-
  Bikes %>% 
  mutate(time_group = round(as.numeric(sdate)/smooth_scale)) %>%
  group_by(time_group) %>%
  summarise(when=min(sdate),count=n())
```

An example ...

```{r}
lungDeaths <- cbind(ldeaths, mdeaths, fdeaths)

dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dySeries("mdeaths", drawPoints = TRUE, color = "blue") %>%
  dySeries("fdeaths", stepPlot = TRUE, color = "red")
```

```{r eval=FALSE}
dygraph(Bikes2, main = "Bike Check-outs") %>%
  dySeries("x", label = "number of bikes") %>% 
  dyRangeSelector()
```

```{r}
Outflux <-
  Bikes %>%
  group_by(sstation, hour) %>%
  summarise(outcount = n())
Influx <- 
  Bikes %>%
  group_by(estation, hour) %>%
  summarize(incount = n())
Both <- inner_join( Outflux, Influx, by=c("sstation"="estation", "hour"="hour"))
```

Bring in latitude and longitude information ...
```{r}
Out_and_influx <-
  Stations %>% 
  select(name, lat, long) %>%
  left_join(Both, by=c("name"="sstation"))
```

```{r}
Out_and_influx %>% 
#  filter(hour == 6) %>%
  ggplot(aes(x=long, y=lat)) +
  geom_point(alpha = 0.2, color="blue", aes(size=outcount)) + 
  geom_point(alpha = 0.2, color="green", aes(size=incount)) + 
  scale_size_area() +
  facet_wrap( ~ hour)
```

Plot by location. facet by time.  Then ... Slider for time.

## Interactive?

What time of day? `r sliderInput("hour", "Hour", min=0, max=23, value=12, step=1)`

```{r  fig.width=9, fig.height=9} 
p <- shiny::reactive( {Out_and_influx %>% 
  filter(hour == input$hour) %>%
  ggplot(aes(x=long, y=lat)) +
  geom_point(alpha = 0.2, color="blue", aes(size=outcount)) + 
  geom_point(alpha = 0.2, color="green", aes(size=incount)) + 
  scale_size_area() +
  ylim(38.85, 38.95) + xlim(-77.06,-76.98)} )
```

```{r}
shiny::renderPlot(p())
```
## Task for DCF Statistics week

Look at distribution of number of uses of various bicycles.


